Local Tuning in Peano Curves-based Global Optimization Scheme

### 2016-07-15 2016/YSC/ysc_paper.tex
Deleted unused header, update YSC paper

The one-dimensional characteristical algorithms represent one of the classes of the
optimization algorithms used widely \cite{optHandbook}. A number of schemes of reduction of the
multidimensional problems to one or several one-dimensional ones, i. e. the \textit{dimension
reduction} schemes have been developed \cite{stronginBook}, \cite{butzCurves}.
\par
In recent years, when building the global optimization algorithms, an increased attention
is paid to accounting for the local properties of the objective function. In the
case of the characteristic schemes, this approach allows developing a mixed algorithm
\cite{mixedAlg} as well as using different estimates of Lipschitz constant in different search
domain \cite{sergLocalTuningFirst}. Originally, the latter approach has been proposed for the one-dimensional
problems, also the investigation of this one when using the nested optimization scheme
has been carried out \cite{nestedLocal}. In all these cases, this approach have demonstrated itself to
be an efficient one, having allowed accelerating the convergence of the
optimization methods essentially. In the present paper, a generalization of the method
for building of the local estimates of Lipschitz constants onto the case of Hölder
functions obtained when reducing the multidimensional optimization problems is considered.

@@

The global optimization problem can be formulated as follows:
to find the global minimum of an \(N\)-dimensional function \(\varphi(y)\) in a hyperinterval
\(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\): \(\varphi(y^*)=\min\{\varphi(y):y\in D\}\).
In order to obtain an estimate of the global minimum from a finite number of computations
of the function values, \(\varphi(y)\) is required to satisfy Lipschitz condition.
\begin{displaymath}
\label{lip}
|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
\end{displaymath}
\par
The use of the evolvents i. e. the curves filling the space is a classic scheme
for the dimension reduction for the global optimization algorithms \cite{strOptBook}.
\begin{displaymath}
\label{cube}
\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
\end{displaymath}
\par
Such a mapping allows reducing a problem stated in a multidimensional space to solving
a one-dimensional one at the expense of worsening its properties. In particular,
the one-dimensional function \(\varphi(y(x))\) is not a Lipschitz one but is a Hölder one:
\begin{displaymath}
\label{holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
\end{displaymath}
where Hölder constant \(H\) is related to Lipschitz constant \(L\) by the relation
\begin{displaymath}
H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}
\end{displaymath}
\par
Therefore, not limiting the generality, one can consider the minimization of the
one-dimensional function \(f(x)=\varphi(y(x))\), \(x\in[0;1]\) satisfying Hölder condition.
The issues of numerical building of the mapping like Peano curve and the corresponding
theory have been considered in details in \cite{strOptBook}. Here we would note that an evolvent
built numerically is an approximation to the theoretical Peano curve with the precision
of the order \(2^{-m}\) where \(m\) is the building parameter of the evolvent.

@@

\begin{theorem}
Assume the objective function \(f(x)\) to  satisfy Hölder condition with finite
constant \(H > 0\), and let x be a limit point of \(\{x_k\}\) generated by the algorithm.
Then, the following assertions hold:
\begin{enumerate}
  \item If \(x\in(0;1)\), then convergence to \(x\) is a bilateral one i. e. there
  exist two infinite subsequences of \(\{x_k\}\) converging to \(x\): one from the
  left, the other from the right;
  \item \(f(x_k) \geqslant f(x)\) for all trial points \(x_k, k \geqslant 1\);
  \item If there exists another limit point \(x^* = x\), then \(f(x) = f(x^*)\);
  \item If the function \(f(x)\) has a finite number of local minima in \([0, 1]\),
  then the point \(x\) is a local optimum one;
  \item (Sufficient conditions for convergence to a global minimizer). Let \(x^*\)
  be a global minimizer of \(f(x)\). If there exists an iteration number \(k^*\)
  such that for all \(k > k^*\) the inequality
  \(M_j(k) > H_j(k)\) holds, where \(H_j(k)\) is Hölder constant for the interval
  \([x_{j(k)-1}, x_{j(k)}]\) containing \(x^*\), and \(M_{j(k)}\) is its estimate.
  Then, the set of limit points of the sequence \(\{xk\}\) coincides with the set
  of global minimizers of the function \(f(x)\).
\end{enumerate}
\end{theorem}

@@

%------------------------------------------------------------------------------
% Refs:
%
\label{sect:bib}
\bibliographystyle{plain}
%\bibliographystyle{alpha}
%\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{ysc_paper}

>>>>

The one-dimensional characteristical algorithms represent one of the classes of the
optimization algorithms used widely \cite{optHandbook}. A number of schemes of reduction of the
multidimensional problems to one or several one-dimensional ones, i. e. the \textit{dimension
reduction} schemes have been developed \cite{stronginBook}, \cite{butzCurves}.
\par
In recent years, when building the global optimization algorithms, an increased attention
is paid to accounting for the local properties of the objective function. In the
case of the characteristical schemes, this approach allows developing a mixed algorithm
\cite{mixedAlg} as well as using different estimates of Lipschitz constant in different search
domain \cite{sergLocalTuningFirst}. Originally, the latter approach has been proposed for the one-dimensional
problems, also the investigation of this one when using the nested optimization scheme
has been carried out \cite{nestedLocal}. In all these cases, this approach have demonstrated itself to
be an efficient one, having allowed accelerating the convergence of the
optimization methods essentially. In the present paper, a generalization of the method
for building of the local estimates of Lipschitz constants onto the case of Hölder
functions obtained when reducing the multidimensional optimization problems is considered.

@@

The global optimization problem can be formulated as follows:
to find the global minimum of an \(N\)-dimensional function \(\varphi(y)\) in a hyperinterval
\(D=\{y\in R^N:a_i\leqslant x_i\leqslant{b_i}, 1\leqslant{i}\leqslant{N}\}\): \(\varphi(y^*)=\min\{\varphi(y):y\in D\}\).
In order to obtain an estimate of the global minimum from a finite number of computations
of the function values, \(\varphi(y)\) is required to satisfy Lipschitz condition.
\begin{displaymath}
\label{lip}
|\varphi(y_1)-\varphi(y_2)|\leqslant L\Vert y_1-y_2\Vert,y_1,y_2\in D,0<L<\infty
\end{displaymath}
\par
The use of the evolvents \(y(x)\) i. e. the curves filling the space is a classic scheme
for the dimension reduction for the global optimization algorithms \cite{strOptBook}.
\begin{displaymath}
\label{cube}
\lbrace y\in R^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\}
\end{displaymath}
\par
Such a mapping allows reducing a problem stated in a multidimensional space to solving
a one-dimensional one at the expense of worsening its properties. In particular,
the one-dimensional function \(\varphi(y(x))\) is not a Lipschitz one but is a Hölder one:
\begin{displaymath}
\label{holder}
|\varphi(y(x_1))-\varphi(y(x_2))|\leqslant H{|x_1-x_2|}^{\frac{1}{N}},x_1,x_2\in[0,1]
\end{displaymath}
where Hölder constant \(H\) is related to Lipschitz constant \(L\) by the relation
\begin{displaymath}
H=4Ld\sqrt{N},d=\max\{b_i-a_i:1\leqslant i\leqslant N\}
\end{displaymath}
\par
Therefore, not limiting the generality, one can consider the minimization of the
one-dimensional function \(f(x)=\varphi(y(x))\), \(x\in[0;1]\) satisfying Hölder condition.
The issues of numerical building of the mapping like Peano curve and the corresponding
theory have been considered in details in \cite{strOptBook}. Here we would note that an evolvent
built numerically is an approximation to the theoretical Peano curve with the precision
of the order \(2^{-m}\) where \(m\) is the building parameter of the evolvent.

@@

\begin{theorem}
Assume the objective function \(f(x)\) to  satisfy Hölder condition with finite
constant \(H > 0\), and let x be a limit point of \(\{x_k\}\) generated by the algorithm.
Then, the following assertions hold:
\begin{enumerate}
  \item If \(x\in(0;1)\), then convergence to \(x\) is a bilateral one i. e. there
  exist two infinite subsequences of \(\{x_k\}\) converging to \(x\): one from the
  left, the other from the right;
  \item \(f(x_k) \geqslant f(x)\) for all trial points \(x_k, k \geqslant 1\);
  \item If there exists another limit point \(x^* = x\), then \(f(x) = f(x^*)\);
  \item If the function \(f(x)\) has a finite number of local minima in \([0, 1]\),
  then the point \(x\) is a local optimum one;
  \item (Sufficient conditions for convergence to a global minimizer). Let \(x^*\)
  be a global minimizer of \(f(x)\). If there exists an iteration number \(k^*\)
  such that for all \(k > k^*\) the inequality
  \(M_j(k) > H_j(k)\) holds, where \(H_j(k)\) is Hölder constant for the interval
  \([x_{j(k)-1}, x_{j(k)}]\) containing \(x^*\), and \(M_{j(k)}\) is its estimate.
  Then, the set of limit points of the sequence \(\{x_k\}\) coincides with the set
  of global minimizers of the function \(f(x)\).
\end{enumerate}
\end{theorem}

@@

%------------------------------------------------------------------------------
% Refs:
%
\label{sect:bib}
%\bibliographystyle{plain}
%\bibliographystyle{alpha}
\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{ysc_paper}
